---
title: Using boto3 (Python) with R (reticulate) to push to S3
author: Matthias Raess
date: '2020-01-20'
slug: python-boto3-reticulate-r
categories:
  - Python
  - Reticulate
  - R
tags:
  - RStudio
keywords:
  - tech
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(python.reticulate = FALSE)
```

## Using Python with R

The motivation for this comes from the fact that it is easier now more than ever to use Python with R via the `reticulate` package.

It does make a lot of sense too! I rely on packages to upload client data to S3 buckets quite a bit using R scripts and `RStudio Connect` for automation. Unfortunately, the `aws.s3` r-package has been wonky and does not seem to be maintained.

Enter, `boto3` - a Python package which lets you interface AWS S3. It is cleaner and easier to handle in terms of the set-up.

## Example

You first set up your R-script like you usually would (assuming you have Python installed and set up) with the addition to the `reticulate` library added.


```{r eval=FALSE, echo=TRUE}

library(tidyverse)
library(reticulate)

## Do stuff and write to file

path <- "./file_name.csv"
s3_path <- "directory/file_name.csv"
  
ACCESS_KEY <- "AWS-KEY"
SECRET_KEY <- "AWS-SECRET"

```

Once you're done processing data, etc. and writing results to a file, you move on to the Python section - you might have even seen the function below in some shape or form before. 

```

# import boto3
# 
# 
# def upload_to_aws(local_file, bucket, s3_file):
#     s3 = boto3.client('s3', aws_access_key_id = r.ACCESS_KEY,
#                       aws_secret_access_key = r.SECRET_KEY)
# 
#     try:
#         s3.upload_file(local_file, bucket, s3_file, ExtraArgs={'ACL':'bucket-owner-full-control'})
#         print("Upload Successful")
#         return True
#     except FileNotFoundError:
#         print("The file was not found")
#         return False
#     except NoCredentialsError:
#         print("Credentials not available")
#         return False
# 
# uploaded = upload_to_aws(r.path, 'bucket_name', r.path_s3)
```

The beauty of this is that you can reference r-objects easily by prepending `r.`, which allows you to not only feed your key/secret pair into the Python script but also the file paths for both the local file and the S3-path.

That's it!!
